---
title: "Milestone 3"
author: "Jacob Hansen"
date: "2/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readxl)
library(readr)
library(janitor)
library(tidyverse)
library(tidycensus)
```

**About**

How can a political party maximize the utility of voter registration? I examine recent voter registration data in Arizona, disaggregated by county, to determine the conditions under which more voter registration best translates to more votes. Using census data, this analysis can be further disaggregated to examine the effects that race, income, and geography have on turning registrations into votes.

```{r, echo=FALSE, message=FALSE, warning=FALSE}

# Because of when this project is being done--early in 2020--I want to maximize
# the data available to me. For this reason, I plan on using January data. I
# plan on using the last few years worth of data, but may decide to reach
# further back if time allows and it seems relevant.

jan_2017 <- read.csv("2017-01-01.csv", skip = 4) %>% 
  clean_names() %>% 
  head(45) %>% 
  select(- x, - x_1, - x_2, - precincts, - total) %>% 
  select(county, date_period, republican, democratic, libertarian, green, other) %>% 
  filter(date_period == "17-Jan") %>% 
  pivot_longer(names_to = "party",
               values_to = "registration_2017",
               cols = c(republican:other))
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
jan_2018 <- read.csv("2018-01-01.csv", skip = 4) %>% 
  clean_names() %>% 
  head(45) %>% 
  select(- x, - x_1, - x_2) %>% 
  select(county, date_period, republican, democratic, libertarian, green, other) %>% 
  filter(date_period == "Jan-18") %>% 
  pivot_longer(names_to = "party",
               values_to = "registration_2018",
               cols = c(republican:other))
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
jan_2019 <- read_csv("2019-01-01.csv", skip = 5) %>% 
  clean_names() %>% 
  head(60) %>% 
  select(- x3, - x5, - x6, - x8, - x9, - x13, - x15, - x17) %>% 
  na.omit() %>% 
  select(county, date_period, republican, democratic, libertarian, green, other) %>% 
  filter(date_period == "Jan-19") %>% 
  pivot_longer(names_to = "party",
               values_to = "registration_2019",
               cols = c(republican:other))
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
jan_2020 <- read_csv("2020-01-21.csv", skip = 5) %>% 
  clean_names() %>% 
  head(60) %>% 
  select(- x3, - x5, - x6, - x8, - x9, - x13, - x15, - x17, - x18, - x19) %>% 
  na.omit() %>% 
  select(county, date_period, republican, democratic, libertarian, green, other) %>% 
  filter(date_period == "Jan-20") %>% 
  pivot_longer(names_to = "party",
               values_to = "registration_2020",
               cols = c(republican:other))
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
jan_2017_to_2018 <- jan_2017 %>% 
  full_join(jan_2018, by = c("county", "party"), suffix = c("_2017", "_2018")) %>% 
  select(- date_period_2017, - date_period_2018)

jan_2017_to_2019 <- jan_2017_to_2018 %>% 
  full_join(jan_2019, by = c("county", "party")) %>% 
  select(- date_period)

jan_2017_to_2020 <- jan_2017_to_2019 %>% 
  full_join(jan_2020, by = c("county", "party")) %>% 
  select(- date_period)
```




